Vertical and oblique cultural transmission fluctuating in time and in space

### 2018-07-13 ms/ms_sunnyvale.tex
first submission to TPB

In evolutionary genetics, the properties of transmission are known to be important, e.g., uniparental vs.\ biparental; haploid vs.\ diploid; with or without recombination; sexual or asexual. For cultural evolution, the mode of transmission between individuals of the same or different generations is also central to the dynamics of cultural traits.
Using analysis with epidemic theory, \citet[ch.~3]{Cavalli-Sforza1981}  introduced models in which a specific cultural trait is transmitted either vertically, that is, directly from parent to offspring, obliquely, that is, from non-parental members of the parent's generation to an offspring, or horizontally, that is, among members of the same generational cohort.
Mathematical models of cultural evolution may include one or more modes of cultural transmission and in most of these models the mode of transmission by individuals is fixed; that is, it does not vary over time {\color{red}or space}.
For example, \cite{Feldman1996}, \cite{Wakano2004}, \cite{Aoki2005}, and \citet{McElreath2008} focused on competition between individual learning, innate learning, and social learning; \citet{Fogarty2017} compared the effects of different oblique mechanisms (random, success-biased, best-of-$k$, one-to-many) on the cultural richness and diversity of the population; and \citet{Aoki2012} modeled scenarios in which individual and social learning occur during separate stages in life.
In these studies, the cultural transmission rule did not change during the evolutionary process.

@@

Here, we analyze a model~(\autoref{fig:transmission}) in which individuals learn from their parents with probability $\rho$ (vertical transmission), and from adults in the parental generation with probability $1-\rho$ (oblique transmission), where $\rho$ fluctuates over time or space.
We find that fluctuations in $\rho$ without corresponding fluctuations in selection cannot maintain a polymorphism.
When both the rate of vertical transmission and selection fluctuate together in a deterministic setting, the stronger the selection, the greater is the difference in vertical transmission rates that maintains a polymorphism.
We also study a two-population model with migration, symmetrically different fitnesses in the two populations, and different rates of vertical transmission in the two populations.
Stability of fixation points  and a polymorphism are shown to depend on strength of selection as well as rates of migration and transmission.

@@

Suppose that both transmission and selection fluctuate together, so that when $A$ is favored, with $w_A=1+s$ and $w_B=1$, the transmission rate is $\rho_A$, and when $B$ is favored, with $w_A=1$ and $w_B=1+s$, the transmission rate is $\rho_B$.
The change in the frequency $x$ of phenotype $A$ when either $A$ or $B$ is favored is described by $F_A(x)$ or $F_B(x)$, respectively, where
\begin{equation}\begin{aligned} \label{eq:recurrence_periodic_fluc}
&F_A(x) = \rho_A \frac{1+s}{1+sx} x + (1-\rho_A)x, &\quad \text{and} \\
&F_B(x) = \rho_B \frac{1}{1+s-sx} x + (1-\rho_B)x.
\end{aligned}\end{equation}

@@

{\bf Result 1.} {\sl Suppose that vertical transmission occurs at rate $\rho_A$ when the fitness of phenotype $A$ is $1+s$ relative to $1$ for phenotype $B$, and at rate $\rho_B$ when the fitness of phenotype $B$ is $1+s$ relative to $1$ for phenotype $A$.
Then the stronger the selection (i.e. the greater the value of $s$), the larger the difference in vertical transmission rates that allows a protected polymorphism.}

@@

% figure
\begin{figure*}[htb]
\centering
\includegraphics[width=0.65\linewidth]{../figures/{rho1_rho2_phase_k=1}.pdf}
\caption{
\textbf{Protected polymorphism.}
The stable equilibrium of the frequency of phenotype $A$ (eq.~\ref{eq:xstar_periodic_fluc_k=1}) for different selection coefficients ($s$ on x-axis) and size of fluctuations in vertical transmission rates ($\rho_B-\rho_A$ on y-axis) when both selection and transmission fluctuate every generation ($k=1$).
Dashed lines represent $\rho_B=\frac{\rho_A}{1+s(1-\rho_A)}$ and $\rho_B=\frac{(1+s)\rho_A}{1+s\rho_A}$, the limits on $\rho_B-\rho_A$ from inequalities~\eqref{eq:poly_condition_periodic_fluc} that permit a protected polymorphism.
Here, $\rho_A=0.5$.}
\label{fig:rho1_rho2_phase_k=1}
\end{figure*}

@@

Now suppose that both selection and transmission fluctuate randomly.
Rewrite eq.~28 from \citet{Ram2018} so that $\rho$ is also a random variable
\begin{equation}
x_{t+1} = x_t \frac{1 + z_t + x_t (1 - \rho_t) s_t}{1 + x_t s_t},
\end{equation}
where $z_t=\rho_t s_t$; $s_t$ are i.i.d (independent and identically distributed); $Pr(-1+C<s_t<D)=1$ for some positive $C$ and $D$; $\rho_t$ are i.i.d; and  $0<\rho_t<1$ ($t=0,1,2,\ldots$).
Therefore, $z_t$ are independent and identically distributed and $P(-1+C < z_t < D)$.

@@

In our case there are two constant equilibria $x^*=0$ and $x^*=1$ corresponding to fixation in $B$ and $A$, respectively.
From \citet[][results~6 and 7]{Ram2018} we can characterize the stochastic local stability of these fixations as follows.

@@

However, note that symmetry of $s_t$ around zero provides an advantage to phenotype $B$: using Jensen's inequality, $E[w_A/w_B] = E[1+s_t] = 1 \le E[1/(1+s_t)] = E[w_B/w_A]$.
Therefore, if we take the i.i.d fitness random variables for $A$ and $B$ to be $w_{A,t}, w_{B,t} \sim U(0,1)$, respectively, and define $s_t=(w_{A,t}-w_{B,t})/w_{B,t}$, then neither $A$ nor $B$ has an advantage, on average (i.e. $E[w_{A,t}/w_{B,t}]=E[w_{B,t}/w_{A,t}]$, see \autoref{fig:beta}A), and both $E[z_t]$ and $E[-z_t/(1+s_t)]$ are positive, so that both fixations are not stochastically locally stable, and we expect the population to approach a polymorphic distribution.

@@

% figure
\begin{figure*}[hbt]
\centering
\includegraphics[width=\linewidth]{../figures/{beta}.png}
\caption{
\textbf{Covariance of selection and transmission.}
\textbf{(A)} Histogram of $w_{A,t}/w_{B,t}$ where $w_{A,t}$ and $w_{B,t}$ are identically and independently distributed uniform random variables $U(0,1)$.
\textbf{(B)} Histogram of $s_t = (w_{A,t}-w_{B,t})/w_{B,t}$.
\textbf{(C)} Histogram of $\rho_t \sim Beta(1+s_t, 1)$.
\textbf{(D)} The joint distribution of $\rho_t$ and $s_t$ demonstrates a positive correlation $cov(s_t, \rho_t)>0$.
}
\label{fig:beta}
\end{figure*}

@@

\autoref{fig:asym_migration} shows some numerical results for the case of symmetric selection{\color{red},} $w_1=1/w_2=w>1$.
We focus on composite parameters of the model:
$m_1/m_2$, on the x-axis, is the ratio of migration rates into deme 1 and deme 2; when this ratio is large, deme 1 accepts more migrants than deme 2.
$\rho_1/\rho_2$, on the y-axis{\color{red},} the ratio of the vertical transmission rates in deme 1 and deme 2; when this ratio is large, individuals in deme 1 use vertical transmission more often then individuals in deme 2.

@@

We have the following results:
\begin{itemize}
\item With only oblique transmission ($\rho_1=\rho_2=0$), there are only neutral equilibria $(x^*,x^*)$ for any $0 \le x^* \le 1$.
\item With only vertical transmission ($\rho_1=\rho_2=1$), the fixation equilibria  $(0,0)$, $(1,1)$ are unstable and there exists a protected polymorphism
\begin{equation}
\begin{aligned} \label{eq:migration_model_unconditional_symmetric_polymorphism}
x_1^* &= \frac{w-1-m(w+1) + \sqrt{\Delta}}{2(w-1)}, \\
x_2^* &= 1-x_1^*,
\end{aligned}
\end{equation}
where $\Delta = m^2(w+1)^2+(1-2m)(w-1)^2$.
\item With only vertical transmission in one deme ($\rho_1=1$) and a combination of both vertical and oblique transmission in the other deme ($\rho_2=\rho$), fixation of $B$ is unstable, and fixation of $A$ is stable if and only if the vertical transmission rate in deme 2 is below $\hat \rho$; that is,
\begin{equation} \label{eq:migration_model_unconditional_symmetric_condition_rho}
\rho < \hat \rho = \frac{m}{m+(1-m)(w-1)} < 1,
\end{equation}
or if the migration rate is above $\hat m$; that is,
\begin{equation} \label{eq:migration_model_unconditional_symmetric_condition_m}
m > \hat m = \frac{\rho w - 1}{\rho w + 1}.
\end{equation}
\end{itemize}
{\color{red}The proofs of (22) and (23) are in Appendix I.}
\paragraph{Examples.}

@@

% figure
\begin{figure*}[ht]
\centering
\includegraphics[width=0.75\linewidth]{../figures/migration_rho.pdf}
\caption{
\textbf{Oblique transmission and migration: effect of transmission.} 
The figure shows $x^*_i$ the stable frequencies of $A$ (top row) and $\overline{w}^*_i$ the {\color{red}stable} population mean fitnesses (bottom row) in the two demes.
Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype $B$ is $1$ in both demes).
The vertical transmission rate is $\rho_1=1$ in deme 1, and $\rho_2$ (x-axis) in deme 2.
Migration rate $m$ is 0.05, 0.1, or 0.25 in the left, middle, and right columns, respectively.
The shaded area denotes stable fixation of phenotype $A$ according to {\color{red}inequality}~\ref{eq:migration_model_unconditional_symmetric_condition_rho}.
Lines are drawn by iterating eqs.~\ref{eq:migration_model_unconditional_symmetric} until frequencies in consecutive generations differ by less than $10^{-4}$, starting with equal frequencies.
}
\label{fig:migration_rho}
\end{figure*}

% figure
\begin{figure*}[ht]
\centering
\includegraphics[width=0.75\linewidth]{../figures/migration_m.pdf}
\caption{
\textbf{Oblique transmission and migration: effect of migration.} 
The figure shows $x^*_i$, the stable frequencies of $A$ (top row), and $\overline{w}^*_i$, the {\color{red}stable} population mean fitnesses (bottom row), in the two demes.
Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype $B$ is $1$ in both demes).
The vertical transmission rate is $\rho_1=1$ in deme 1, and $\rho_2=1$, $0.4$, and $0.2$, in the left, middle, and right columns, respectively, in deme 2.
Migration rate $m$ is on the x-axis.
The shaded area denotes stables fixation of phenotype $A$ according to {\color{red}inequality}~\ref{eq:migration_model_unconditional_symmetric_condition_m}.
Lines are drawn by iterating eq.~\ref{eq:migration_model_unconditional_symmetric} until frequencies in consecutive generations differ by less than $10^{-4}$, starting with equal frequencies.
}
\label{fig:migration_m}
\end{figure*}

@@

One caveat of our model is the use of the ``phenotypic gambit''~\citep{Grafen1984}: the assumption that the transmission mode itself is strictly vertically transmitted.
Although there is some evidence that the tendency to use different learning mechanisms is genetically transmitted~\citep{Foucaud2013}, this assumption can be challenged: individuals may be able to learn how and when to learn, in what has been called ``social learning of social learning''~\citep{Mesoudi2016}.
Indeed, it has been demonstrated that guppy fish are more likely to learn from others if previous social experiences provided benefits~\citep{Leris2016}.
It is also possible that the transmission mode is regulated.
For example, \citet{Farine2015} found that zebra fish switch from vertical to oblique transmission after exposure to stress hormones.
Our model accounts for cases in which the entire population changes its transmission mode due to stress, but not for cases in which only {\color{red}specific (e.g. stressed)} individuals do so.

@@

First, the discriminant $\Delta$ of $f_1(x)=0$ is $b^2-4ac$ for $m=0$ and $m=1/2$ is
\begin{equation}\begin{aligned}
&\Delta(m=0) = \Big(\frac{(w \rho + 1) (w-1)}{w}\Big)^2, \quad \text{and} \\
&\Delta(m=1/2) = \Big(\frac{w^2 \rho + w(1-\rho) + 1}{2 w}\Big)^2,
\end{aligned}\end{equation}
which are both positive. 
Writing the discriminant $\Delta$ of $f_1(x)$ as a polynomial in $m$, $g(m)$, the discriminant of $g(m)$ is
$$
\big(- 16 (w\rho + 1)^2 (w - 1)^2 \big(\rho w + 1 - \rho \big) / w^3\big),
$$
which is negative, and therefore $\Delta$ is positive for any $0 \le m \le 1/2$ and $f_1(x)$ has two real roots. \\
Second, $a>0$, $b<0$, and $c>0$ so $f_1$ is positive with a negative derivative at $x=0$ and  a positive derivative at infinity.
Therefore, fixation of $A$ is stable {\color{red}if and only if} both $f_1(1)=a+b+c$ and $f_1'(1)=2a+b$ are positive.
{\color{red}Now $f_1(1)>0$  if $\rho < \frac{m}{m+(1-m)(w-1)}=M_1$, and
$f_1'(1)>0$  if $\rho < \frac{m(w+1)+w-1}{(1-m)w(w-1)}=M_2$.
The difference $M_1-M_2$ is a quadratic in $w$ with a negative discriminant $2m-1$ and negative value $-1/2$ at $(w=2, m=0)$, so $M_1<M_2$ for any $w>1$ and $0 \le m \le 1/2$.}

Therefore, fixation of $A$ is stable if {\color{red}$\rho < M_1 = \frac{m}{m+(1-m)(w-1)}$}~\eqref{eq:migration_model_unconditional_symmetric_condition_rho}. By rearranging this inequality, we can also obtain an expression for $m$~\eqref{eq:migration_model_unconditional_symmetric_condition_m}.

>>>>

In evolutionary genetics, the properties of transmission are known to be important, e.g., uniparental vs.\ biparental; haploid vs.\ diploid; with or without recombination; sexual or asexual. For cultural evolution, the mode of transmission between individuals of the same or different generations is also central to the dynamics of cultural traits.
Using analysis with epidemic theory, \citet[ch.~3]{Cavalli-Sforza1981}  introduced models in which a specific cultural trait is transmitted either vertically, that is, directly from parent to offspring, obliquely, that is, from non-parental members of the parent's generation to an offspring, or horizontally, that is, among members of the same generational cohort.
Mathematical models of cultural evolution may include one or more modes of cultural transmission and in most of these models the mode of transmission by individuals is fixed; that is, it does not vary over time or space.
For example, \cite{Feldman1996}, \cite{Wakano2004}, \cite{Aoki2005}, and \citet{McElreath2008} focused on competition between individual learning, innate learning, and social learning; \citet{Fogarty2017} compared the effects of different oblique mechanisms (random, success-biased, best-of-$k$, one-to-many) on the cultural richness and diversity of the population; and \citet{Aoki2012} modeled scenarios in which individual and social learning occur during separate stages in life.
In these studies, the cultural transmission rule did not change during the evolutionary process.

@@

Here, we analyze a model~(\autoref{fig:transmission}) in which individuals learn from their parents with probability $\rho$ (vertical transmission), and from adults in the parental generation with probability $1-\rho$ (oblique transmission), where $\rho$ fluctuates over time or space.
We find that fluctuations in $\rho$ without corresponding fluctuations in selection cannot maintain a polymorphism.
When both the rate of vertical transmission and selection fluctuate together in a deterministic setting, the stronger the selection, the greater is the difference in vertical transmission rates that maintains a polymorphism.
Versions of the model with random fitnesses and transmission rates, and with random drift due to finite population size are also analyzed. We also study a two-population model with migration, symmetrically different fitnesses in the two populations, and different rates of vertical transmission in the two populations.
Stability of fixation points  and a polymorphism are shown to depend on strength of selection as well as rates of migration and transmission.

@@

Suppose that both transmission and selection fluctuate together, so that when $A$ is favored, with $w_A=1+s$ and $w_B=1$ with $s>0$, the transmission rate is $\rho_A$, and when $B$ is favored, with $w_A=1$ and $w_B=1+s$, the transmission rate is $\rho_B$.
The change in the frequency $x$ of phenotype $A$ when either $A$ or $B$ is favored is described by $F_A(x)$ or $F_B(x)$, respectively, where
\begin{equation}\begin{aligned} \label{eq:recurrence_periodic_fluc}
&F_A(x) = \rho_A \frac{1+s}{1+sx} x + (1-\rho_A)x, &\quad \text{and} \\
&F_B(x) = \rho_B \frac{1}{1+s-sx} x + (1-\rho_B)x.
\end{aligned}\end{equation}

@@

{\bf Result 1.} {\sl Suppose that vertical transmission occurs at rate $\rho_A$ when the fitness of phenotype $A$ is $1+s$ relative to $1$ for phenotype $B$, and at rate $\rho_B$ when the fitness of phenotype $B$ is $1+s$ relative to $1$ for phenotype $A$, where $s>0$.
Then the stronger the selection (i.e. the greater the value of $s$), the larger the difference in vertical transmission rates that allows a protected polymorphism.}

@@

% figure
\begin{figure*}[htb]
\centering
\includegraphics[width=0.65\linewidth]{../figures/{rho1_rho2_phase_k=1}.pdf}
\caption{
\textbf{Protected polymorphism.}
The frequency of phenotype $A$ at the stable equilibrium   (eq.~\ref{eq:xstar_periodic_fluc_k=1}) for different selection coefficients ($s$ on x-axis) and size of fluctuations in vertical transmission rates ($\rho_B-\rho_A$ on y-axis) when both selection and transmission fluctuate every generation ($k=1$).
Dashed lines represent $\rho_B=\frac{\rho_A}{1+s(1-\rho_A)}$ and $\rho_B=\frac{(1+s)\rho_A}{1+s\rho_A}$, the limits on $\rho_B-\rho_A$ from inequalities~\eqref{eq:poly_condition_periodic_fluc} that permit a protected polymorphism.
Here, $\rho_A=0.5$.}
\label{fig:rho1_rho2_phase_k=1}
\end{figure*}

@@

Now suppose that both selection and transmission fluctuate randomly.
Rewrite eq.~28 from \citet{Ram2018} so that $\rho$ is also a random variable
\begin{equation}
x_{t+1} = x_t \frac{1 + z_t + x_t (1 - \rho_t) s_t}{1 + x_t s_t},
\end{equation}
where $z_t=\rho_t s_t$; $s_t$ are i.i.d (independent and identically distributed); $Pr(-1+C<s_t<D)=1$ for some positive $C$ and $D$; $\rho_t$ are i.i.d; and  $0<\rho_t<1$ ($t=0,1,2,\ldots$).
Therefore, $z_t$ are independent and identically distributed and $P(-1+C < z_t < D)$.  Following \citet{Karlin1974}, we have:

@@

In our case there are two constant equilibria $x^*=0$ and $x^*=1$ corresponding to fixation in $B$ and in $A$, respectively.
From \citet[][results~6 and 7]{Ram2018} we can characterize the stochastic local stability of these fixations as follows.

@@

However, note that symmetry of $s_t$ around zero provides an advantage to phenotype $B$: using Jensen's inequality, $E[w_A/w_B] = E[1+s_t] = 1 \le E[1/(1+s_t)] = E[w_B/w_A]$.
Therefore, if we take the i.i.d fitness random variables for $A$ and $B$ to be $w_{A,t}, w_{B,t} \sim U(0,1)$, respectively, and define $s_t=(w_{A,t}-w_{B,t})/w_{B,t}$, then neither $A$ nor $B$ has an advantage, on average (i.e. $E[w_{A,t}/w_{B,t}]=E[w_{B,t}/w_{A,t}]$, see \autoref{fig:beta}a), and both $E[z_t]$ and $E[-z_t/(1+s_t)]$ are positive, so that both fixations are not stochastically locally stable, and we expect the population to approach a polymorphic distribution.

@@

% figure
\begin{figure*}[hbt]
\centering
\includegraphics[width=\linewidth]{../figures/{beta}.pdf}
\caption{
\textbf{Covariance of selection and transmission.}
\textbf{(a)} Histogram of $w_{A,t}/w_{B,t}$ where $w_{A,t}$ and $w_{B,t}$ are identically and independently distributed uniform random variables $U(0,1)$.
\textbf{(b)} Histogram of $s_t = (w_{A,t}-w_{B,t})/w_{B,t}$.
\textbf{(c)} Histogram of $\rho_t \sim Beta(1+s_t, 1)$.
\textbf{(d)} The joint distribution of $\rho_t$ and $s_t$ demonstrates a positive correlation $cov(s_t, \rho_t)>0$.
}
\label{fig:beta}
\end{figure*}

@@

\autoref{fig:asym_migration} shows some numerical results for the case of symmetric selection, $w_1=1/w_2=w>1$.
We focus on composite parameters of the model:
$m_1/m_2$, on the x-axis, is the ratio of migration rates into deme 1 and deme 2; when this ratio is large, deme 1 accepts more migrants than deme 2.
$\rho_1/\rho_2$ on the y-axis is the ratio of the vertical transmission rates in deme 1 and deme 2; when this ratio is large, individuals in deme 1 use vertical transmission more often then individuals in deme 2.

@@

We have the following results:
\begin{itemize}
\item With only oblique transmission ($\rho_1=\rho_2=0$), there are only neutral equilibria $(x^*,x^*)$ for any $0 \le x^* \le 1$.
\item With only vertical transmission ($\rho_1=\rho_2=1$), the fixation equilibria  $(0,0)$, $(1,1)$ are unstable and there exists a protected polymorphism
\begin{equation}
\begin{aligned} \label{eq:migration_model_unconditional_symmetric_polymorphism}
x_1^* &= \frac{w-1-m(w+1) + \sqrt{\Delta}}{2(w-1)}, \\
x_2^* &= 1-x_1^*,
\end{aligned}
\end{equation}
where $\Delta = m^2(w+1)^2+(1-2m)(w-1)^2$.
\item With only vertical transmission in one deme ($\rho_1=1$) and a combination of both vertical and oblique transmission in the other deme ($\rho_2=\rho$), fixation of $B$ is unstable, and fixation of $A$ is stable if and only if the vertical transmission rate in deme 2 is below $\hat \rho$; that is,
\begin{equation} \label{eq:migration_model_unconditional_symmetric_condition_rho}
\rho < \hat \rho = \frac{m}{m+(1-m)(w-1)} < 1,
\end{equation}
or if the migration rate is above $\hat m$; that is,
\begin{equation} \label{eq:migration_model_unconditional_symmetric_condition_m}
m > \hat m = \frac{\rho w - 1}{\rho w + 1}.
\end{equation}
\end{itemize}
The proofs of (22) and (23) are in Appendix I.
\paragraph{Examples.}

@@

% figure
\begin{figure*}[ht]
\centering
\includegraphics[width=0.75\linewidth]{../figures/migration_rho.pdf}
\caption{
\textbf{Oblique transmission and migration: effect of transmission.} 
The figure shows $x^*_i$ the stable frequencies of $A$ (top row) and $\overline{w}^*_i$ the stable population mean fitnesses (bottom row) in the two demes.
Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype $B$ is $1$ in both demes).
The vertical transmission rate is $\rho_1=1$ in deme 1, and $\rho_2$ (x-axis) in deme 2.
Migration rate $m$ is 0.05, 0.1, or 0.25 in the left, middle, and right columns, respectively.
The shaded area denotes stable fixation of phenotype $A$ according to inequality~\ref{eq:migration_model_unconditional_symmetric_condition_rho}.
Lines are drawn by iterating eqs.~\ref{eq:migration_model_unconditional_symmetric} until frequencies in consecutive generations differ by less than $10^{-4}$, starting with equal frequencies.
}
\label{fig:migration_rho}
\end{figure*}

% figure
\begin{figure*}[ht]
\centering
\includegraphics[width=0.75\linewidth]{../figures/migration_m.pdf}
\caption{
\textbf{Oblique transmission and migration: effect of migration.} 
The figure shows $x^*_i$, the stable frequencies of $A$ (top row), and $\overline{w}^*_i$, the stable population mean fitnesses (bottom row), in the two demes.
Selection is symmetric between the two demes (the fitness of phenotype $A$ is $w_1=2$ in deme 1 and $w_2=0.5$ in deme 2; the fitness of phenotype $B$ is $1$ in both demes).
The vertical transmission rate is $\rho_1=1$ in deme 1, and $\rho_2=1$, $0.4$, and $0.2$, in the left, middle, and right columns, respectively, in deme 2.
Migration rate $m$ is on the x-axis.
The shaded area denotes stables fixation of phenotype $A$ according to inequality~\ref{eq:migration_model_unconditional_symmetric_condition_m}.
Lines are drawn by iterating eq.~\ref{eq:migration_model_unconditional_symmetric} until frequencies in consecutive generations differ by less than $10^{-4}$, starting with equal frequencies.
}
\label{fig:migration_m}
\end{figure*}

@@

One caveat of our model is the use of the ``phenotypic gambit''~\citep{Grafen1984}: the assumption that the transmission mode itself is strictly vertically transmitted.
Although there is some evidence that the tendency to use different learning mechanisms is genetically transmitted~\citep{Foucaud2013}, this assumption can be challenged: individuals may be able to learn how and when to learn, in what has been called ``social learning of social learning''~\citep{Mesoudi2016}.
Indeed, it has been demonstrated that guppy fish are more likely to learn from others if previous social experiences provided benefits~\citep{Leris2016}.
It is also possible that the transmission mode is regulated.
For example, \citet{Farine2015} found that zebra fish switch from vertical to oblique transmission after exposure to stress hormones.
Our model accounts for cases in which the entire population changes its transmission mode due to stress, but not for cases in which only specific (e.g. stressed) individuals do so.

@@

First, the discriminant $\Delta$ of $f_1(x)=0$ is $b^2-4ac$ for $m=0$ and $m=1/2$ is
\begin{equation}\begin{aligned}
&\Delta(m=0) = \Big(\frac{(w \rho + 1) (w-1)}{w}\Big)^2, \quad \text{and} \\
&\Delta(m=1/2) = \Big(\frac{w^2 \rho + w(1-\rho) + 1}{2 w}\Big)^2,
\end{aligned}\end{equation}
which are both positive. 
Writing the discriminant $\Delta$ of $f_1(x)$ as a polynomial in $m$, $g(m)$, the discriminant of $g(m)$ is
$$
\big(- 16 (w\rho + 1)^2 (w - 1)^2 \big(\rho w + 1 - \rho \big) / w^3\big),
$$
which is negative, and therefore $\Delta$ is positive for any $0 \le m \le 1/2$ and $f_1(x)$ has two real roots. \\
Second, $a>0$, $b<0$, and $c>0$ so $f_1$ is positive with a negative derivative at $x=0$ and  a positive derivative at infinity.
Therefore, fixation of $A$ is stable if and only if both $f_1(1)=a+b+c$ and $f_1'(1)=2a+b$ are positive.
Now $f_1(1)>0$  if $\rho < \frac{m}{m+(1-m)(w-1)}=M_1$, and
$f_1'(1)>0$  if $\rho < \frac{m(w+1)+w-1}{(1-m)w(w-1)}=M_2$.
The difference $M_1-M_2$ is a quadratic in $w$ with a negative discriminant $2m-1$ and negative value $-1/2$ at $(w=2, m=0)$, so $M_1<M_2$ for any $w>1$ and $0 \le m \le 1/2$.

Therefore, fixation of $A$ is stable if $\rho < M_1 = \frac{m}{m+(1-m)(w-1)}$~\eqref{eq:migration_model_unconditional_symmetric_condition_rho}. By rearranging this inequality, we can also obtain an expression for $m$~\eqref{eq:migration_model_unconditional_symmetric_condition_m}.
